---
title: "Exam One"
format: html
---

# Load Packages

```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(maps)
```

# Get Data

```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption 
# The first line loads the data from Tidy Tuesday into R. 
# Then, the second line pulls the food consumption data set and stores it into 'fc'.  
```

# Understand Data

```{r}
dim(fc)
str(fc)
head(fc)
```
# Observations 

- The units of observation are individual categories of food within separate countries. 
- There are 11 food categories. 
- There are 130 countries. 

# Understand 
*Grand Research Question*
What does the consumption of each food category in each country look like? 
- This question involves 3 variables (one numerical, two categorical). 

*Country* 
```{r, fig.cap = "Country Visualization"}
#| fig-height: 13
#| fig-width: 15

ggplot(fc, aes(y = country)) + 
  geom_bar(color = "violetred", fill = "orchid2") + 
  labs(y = "Country", x = "Count", title = "Country Distribution")
```
- Each country has a count of 11 because there are 11 food categories. 


*Food Category* 
```{r, fig.cap = "Food Category Visualization"}
#| fig-height: 9
#| fig-width: 15

ggplot(fc, aes(y = food_category)) + 
  geom_bar(fill = "slateblue", color = "royalblue4") + 
  labs(x = "Count", y = "Food Category", title = "Food Category Distribution")
```
- Since there are 130 countries, each category of food has a count of 130. 


*Consumption* 
```{r, fig.cap = "Consumption Visualization"}
#| fig-height: 5
#| fig-width: 10

ggplot(fc, aes(x = consumption)) + 
  geom_density(color = "violetred", fill = "orchid2") + 
  labs(x = "Consumption", y = "Density (kg/person/year)", title = "Consumption Distribution")
```
- Consumption tends to be around 0 to 50 kg/person/year, but some observations get to as high as 450 kg/person/year. It is hard to say much without breaking it down by country, food category, or both. 

*Consumption + Category* 
```{r, fig.cap = "Consumption and Food Category Visualization"}
#| fig-height: 12
#| fig-width: 16

ggplot(fc, aes(x = consumption, fill = food_category)) +
  geom_histogram() +
  facet_wrap(~food_category) +
  scale_fill_viridis_d() + 
  labs(x = "Consumption (kg/person/year)", y = "Density", fill = "Food Category", title = "Consumption by Food Category") 
```

- It appears that the soybean category is on the lower end of consumption, with many observations close to or equaling zero. 
- The milk category has the largest range, reaching all the way to around 450 kg/person/year. Wheat/wheat products and rice also have a relatively wide range, reaching a consumption of about 200. Fish has an outlier near 200 as well. 
- Most of the categories seem to stay in the range of about 0 to 50 kg/person/year. 


*Consumption + Country* 
```{r, fig.cap = "Consumption and Country Visualization"}
#| fig-height: 12
#| fig-width: 23

ggplot(fc, aes(x = consumption, fill = country)) +
  geom_histogram() +
  facet_wrap(~country) +
  scale_fill_viridis_d() + 
  labs(x = "Consumption (kg/person/year)", y = "Density", fill = "Country", title = "Consumption by Country") 
```

- All countries are right-skewed in their consumption, but some have lower peaks and a longer range than others, such as the USA, Israel, Slovenia, and others. Overall, most observations across countries tend to be on the lower end of consumption. 

# Answering Grand RQ 

```{r, fig.cap = "Grand Research Question"}
#| fig-height: 20
#| fig-width: 23

ggplot(fc, aes(x = food_category, y = country, fill = consumption)) +
  geom_tile() +
  scale_fill_viridis_c() + 
  labs(x = "Food Category", y = "Country", fill = "Consumption (kg/person/year)", title = "Consumption by Country and Food Category") +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
```
- There are clear trends that are consistent in most of the countries shown. Milk (including cheese) products are consumed the most overall, with most countries reaching at least 200 kg/person/year and some even higher. Finland notably uses more than 400 kg/person/year. 
- The second most popular food category is wheat and wheat products. They seem to range between 100 and 200 kg/person/year. Tunisia seems to use this category the most, at about 200 kg/person/year. 
- Wheat and wheat products is followed by rice. Bangladesh seems to use it the most at a little lower than 200 kg/person/year.
- Soybeans is used the least out of all of the food categories. It maybe reaches around 25 or so kg/person/year, but mostly stays on the lowest end of the spectrum. 
- All of the other food categories are relatively similar. Lamb, goat, and nuts (including peanut butter) are on the lower end, but they are all stay around 0 to 50 kg/person/year, with few exceptions (fish reaches about 200 in Maldives). 


