---
title: "Notes"
---
# Solo Project Notes 

Need 
- *geometry variable* - describing boundary of areas wanting to color 
  - ggplot2
- *variable* to color areas based on its values 
  - sf 
  
Example: 
Vietnam 
1) Search natural earth package and find Vietnam shapefile 
2) Find package associated with Vietnam 

On the assignment page for solo project is where to find potential datasets. 
There's a video to help describe the sf package and plotting cloropleth maps. 

# 4.1 Review - Bivariate Viz 

```{r}
# Import data
survey <- read.csv("https://ajohns24.github.io/data/112/about_us_2024.csv")

# How many students have now filled out the survey? 
dim(survey)
#28 students. 

# What type of variables do we have? 
str(survey)
#Character, integer, numeric, and another character. 
```
**Example 1**: Hangout Preferences 
```{r}
library(tidyverse)

ggplot(survey, aes(x = hangout)) + 
  geom_bar(color = "white", fill = "pink") + 
  labs(x = "Hangout Spot", y = "Number of Students")
```

There was some variability in answers - most students preferred a beach as a hangout spot, and the least amount of students favored a prairie. 

**Example 2**: Temperature Preferences

```{r}
#Plot 1 
ggplot(survey, aes(x = fave_temp)) + 
  geom_histogram(color = "white", binwidth = 15) + 
  labs(x = "Favorite Temperature (Celsius)", y = "Number of Students")

#Plot 2 
ggplot(survey, aes(x = fave_temp)) + 
  geom_density(color = "purple") + 
  labs(x = "Favorite Temperature (Celsius)", y = "Number of Students")
```
Responses were typically around 20 degrees Celsius, with responses typically in the range of 0 degrees to 30 degrees. There is an outlier at about 85 degrees. 

# 4.2 New Stuff 
**Exploring relationships** 
```{r}
weather <- data.frame(temp_3pm = c(24, 26, 20, 15, 15, 0, 40, 60, 57, 44, 51, 75),
                      location = rep(c("A", "B"), each = 6))

weather

ggplot(weather, aes(x = temp_3pm)) +
      geom_density()
```
```{r}
weather <- data.frame(rain_today = c("no", "no", "no", "no", "yes", "no", "yes", "no", "yes", "yes", "no", "yes"),
                        location = c(rep("A", 7), rep("B", 5)))
    weather
    
ggplot(weather, aes(x = location)) +
      geom_bar()
```
# 5.1 Review
```{r}
library(tidyverse)

# Import data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

# Check out the first 6 rows
head(weather)

# What are the units of observation? -> 2367
nrow(weather)

# How many data points do we have? -> 2367
dim(weather)

# What type of variables do we have?
str(weather)
#We have integer, numeric, and character variables in this data set. 
```
**Example 1** 
```{r}
ggplot(weather, aes (x = temp3pm)) + 
  geom_density()
```
**Example 2** 
Construct 3 plots that address the following research question:

How do afternoon temperatures (temp3pm) differ by location?

```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5)
```
```{r}
# Plot 2 (no facets or densities)
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_histogram()
```

```{r}
# Plot 3 (facets)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density(alpha = 0.5) + 
  facet_wrap(~location)
```

**Example 3** 
```{r}
# Don't worry about the syntax (we'll learn it soon)
woll <- weather |>
  filter(location == "Wollongong") |> 
  mutate(date = as.Date(date))  
```

```{r}
# How often does it raintoday?
# Fill your geometric layer with the color blue.
ggplot(woll, aes(x = raintoday)) + 
  geom_density(fill = "blue")
```
(COME BACK TO)




##Homework 6 Opt Exercises for Quiz Studying

### Optional Exercise: funding time

NOTE: This exercise won't be graded, but it is strongly recommended as additional practice for the quiz.

How many days does it take borrowers to get a loan funded? Let's explore the `days_to_fund` variable in the `loans_2`.

#### Part a

Construct a univariate visualization of `days_to_fund`.

```{r}

```

#### Part b

Construct a plot of `days_to_fund` (y-axis) vs `funded_amount` (x-axis) for each loan in `loans_2`.

```{r}

```

#### Part c

Construct a plot of `days_to_fund` (y-axis) vs `funded_amount` (x-axis) in each `countries.region`. Don't represent each loan on this plot. Rather, include curves to represent the trends in each region.

```{r}

```

#### Part d

Summarize, in words, some takeaway messages from parts a-c about how long it takes to get a loan funded.



### Optional Exercise: loan status

NOTE: This exercise won't be graded, but it is highly recommended as additional practice for the quiz.

The `status` variable in `loans_2` indicates the status of a loan:

```{r}
# Check out a table summary after you've defined loans_2
# loans_2 %>% 
#   count(status)
```

Let's focus here on the patterns in loans that were paid back and those that were defaulted, i.e. not paid back.

#### Part a

Define a new dataset, `defaults`, that only includes loans that were either defaulted or paid.

```{r}
# Define defaults

# Confirm that defaults has 7010 rows and 13 columns

```

#### Part b

Using `defaults`, construct a visualization of the relationship between the funded amount and status of a loan.

```{r}

```

#### Part c

Define a new dataset with four columns:

-   `partner_id`
-   number of defaulted loans through that partner
-   number of loans completely paid back through that partner
-   percentage of loans defaulted

Sort your table from highest default percentage to lowest, and print out only those with at least a 50% default percentage. HINT: You'll have to reshape the data in this process.

```{r}

```

#### Part d

Provide some take-away messages about loan defaults using your results in parts a-d.


