---
title: "Spatial Viz"
---

**Load** 
```{r}
library(tidyverse)
```

## 6.4 Exercises 

# Exercise 1: A leaflet with markers / points 
```{r}
fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

# Check it out
head(fave_places)
```
**Part a** 

```{r}
# Load the leaflet package
library(leaflet)
```

```{r}
# Just a plotting frame
leaflet(data = fave_places)
```

```{r}
# Now what do we have?
leaflet(data = fave_places) |> 
  addTiles()
```

```{r}
# Now what do we have?
# longitude and latitude refer to the variables in our data
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)
```

```{r}
# Since we named them "longitude" and "latitude", the function
# automatically recognizes these variables. No need to write them!
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers()
```

# Exercise 2: Details  
```{r}
# Load package needed to change color
library(gplots)

# We can add colored circles instead of markers at each location
leaflet(data = fave_places) |> 
  addTiles() |> 
  addCircles(color = col2hex("red"))
```
```{r}
# We can change the background
# Mark locations with yellow dots
# And connect the dots, in their order in the dataset, with green lines
# (These green lines don't mean anything here, but would if this were somebody's travel path!)
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |>
  addCircles(weight = 10, opacity = 1, color = col2hex("yellow")) |>
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = col2hex("green")
  )
```

# Exercise 3: Your turn 

```{r}
# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")
```

```{r}
# Don't worry about the syntax
starbucks_mn <- starbucks |>   
  filter(Country == "US", State.Province == "MN")

leaflet(data = starbucks_mn) |>
  addTiles() |>
  addMarkers()
```

## Part 2: Static points on a map 
Learning to use ggplot() instead of leaflet(). 

# Exercise 3: A simple scatterplot 

```{r}
ggplot(starbucks, aes(x = Latitude, y = Longitude)) + 
  geom_point(alpha = 0.2, small = 0.2)
```
# Exercise 4: Adding a country-level background 
Let's add a background map of country-level boundaries. 

**Part a** 
```{r}
# Load the package
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")
```

**Part b**

```{r}
# What does this code produce?
# What geom are we using for the point map?
ggplot(world_boundaries) + 
  geom_sf()
```
```{r}
# Load package needed to change map theme
library(mosaic)

# Add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen"
  ) +
  theme_map()
```
**Part c** 
Starbucks is largely popular in the United States and the majority of Europe. The eastern parts of Asia also seem to have a number of locations as well. 

# Exercise 5: Zooming in on some countries 

**Part a** 
```{r}
# We'll learn this syntax soon! Don't worry about it now.
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))
```

```{r}
cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")
```

**Part b**: Make the map! 

```{r}
# Just the boundaries
ggplot(cma_boundaries) + 
  geom_sf()
```
```{r}
# Add the points
# And zoom in
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50)) +
  theme_map()
```
# Exercise 6: A state and county-level map 

**Part a** 
Need two pieces of info to make this map: 
  1. Data on Starbucks for only the states of interest:  
```{r}
starbucks_midwest <- starbucks |> 
  filter(State.Province %in% c("MN", "ND", "SD", "WI"))
```
  2. A background map of state- and county- level boundaries in these states. 
```{r}
# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)
```
**Part b** 

```{r}
ggplot(midwest_boundaries) + 
   geom_sf() + 
   geom_point(
     data = starbucks_midwest,
     aes(x = Longitude, y = Latitude),
     alpha = 0.7,
     size = 0.2, 
     color = 'darkgreen'
   ) + 
   theme_map()
```
# Exercise 7: Contour maps 

```{r}
# What changed in the plot?
# What changed in our code?!
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
#Instead of geom_point(), we are now using geom_density_2d. 
```
## Part 3: Choropleth maps 

Spatial data isn't always in the form of point locations - ex: the state and county-level data on presidential elections. 
```{r}
elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

```{r}
# Don't worry about the code!

elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))
```

# Exercise 8: State-level choropleth maps 
Let's map the 2020 Republican support in each *state*, repub_pct_20. 

**Part a** 
Need 2 pieces of info: 
  1. Data on elections in each state: elections_by_state 
  2. A background map of state boundaries in the US. 

```{r}
# Get the latitude and longitude coordinates of state boundaries
states_map <- map_data("state")

# Check it out
head(states_map)
```
**Important detail:** Note that the region variable in states_map, and the state_name variable in elections_by_state both label states by the full name in lower case letters. This is critical to the background map and our data being able to communicate.

```{r}
head(states_map)
```
```{r}
head(elections_by_state) 
```
**Part b** - map repub_pct_20 by state

```{r}
# Note where the dataset, elections_by_state, is used
# Note where the background map, states_map, is used
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() 
```
```{r}
# Make it nicer!
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_gradientn(name = "% Republican", colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5)))
```
It’s not easy to get fine control over the color scale for the quantitative repub_pct_20 variable. Instead, let’s plot the discretized version, repub_20_categories:
```{r}
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()
```

```{r}
# Load package needed for refining color palette
library(RColorBrewer)

# Now fix the colors
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```
**Part c** 

```{r}
# Get only the starbucks data from the US
starbucks_us <- starbucks |> 
  filter(Country == "US")

# Map it
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  geom_point(
    data = starbucks_us,
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```
*Details*: 
- map_id is a required aesthetic for geom_map().
  - It specifies which variable in our dataset indicates the region (here state_name).
  - It connects this variable (state_name) to the region variable in our mapping background (states_map). These variables must have the same possible outcomes in order to be matched up (alabama, alaska, arizona,…).
- expand_limits() assures that the map covers the entire area it’s supposed to, by pulling longitudes and latitudes from the states_map.

**Part d** 
We use geom_map() instead of geom_sf() for choropleth maps. (sf is for point maps)

# Exercise 9: County-level choropleth maps 
Time to map the 2020 Republican support in each *county*.

**Part a** 
Need two pieces of info: 
  1. Data on elections in each county (elections_by_county)
  2. A background map of county, boundaries in the US, stored in the county_map dataset in the socviz package: 

```{r}
# Get the latitude and longitude coordinates of county boundaries
library(socviz)
data(county_map) 

# Check it out
head(county_map)
```
```{r}
# Add 0's at the beginning of any fips_code that's fewer than 5 numbers long
# Don't worry about the syntax
elections_by_counties <- elections_by_counties |> 
  mutate(county_fips = as.character(county_fips)) |> 
  mutate(county_fips = 
           ifelse(nchar(county_fips) == 4, paste0("0", county_fips), county_fips))
```

**Part b**

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = repub_20_categories)) +
  geom_map(map = county_map) +
  scale_fill_manual(values = rev(brewer.pal(10, "RdBu")), name = "% Republican") +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal()
```
**Exercise 10: Play around!**

```{r}
ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_rent)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal() + 
  scale_fill_gradientn(name = "median rent", colors = c("white", "plum", "purple"))
```

