---
title: "Strings"
number-sections: true
execute: 
  warning: false
fig-env: 'figure'
fig-pos: 'h'
fig-align: center
code-fold: false
---


## Strings

**Strings**
    When working with character strings, we might want to detect, replace, or extract certain *patterns*. 

### Essential Functions {-}

The `stringr` package within `tidyverse` contains lots of functions to help process strings. We'll focus on the most common. Letting `x` be a string variable...

| function            | arguments                 | returns           |
|:--------------------|:--------------------------|:------------------|
| `str_replace()`     | `x, pattern, replacement` | a modified string |
| `str_replace_all()` | `x, pattern, replacement` | a modified string |
| `str_to_lower()`    | `x`                       | a modified string |
| `str_sub()`         | `x, start, end`           | a modified string |
| `str_length()`      | `x`                       | a number          |
| `str_detect()`      | `x, pattern`              | TRUE/FALSE        |

### Example 1 {-}

Consider the following data with string variables :

```{r}
library(tidyverse)

classes <- data.frame(
  sem        = c("SP2023", "FA2023", "SP2024"),
  area       = c("History", "Math", "Anthro"),
  enroll     = c("30 - people", "20 - people", "25 - people"),
  instructor = c("Ernesto Capello", "Lori Ziegelmeier", "Arjun Guneratne")
)

classes
```

Using only your intuition, use our `str_` functions to complete the following. NOTE: You might be able to use other wrangling verbs in some cases, but focus on the new functions here.

```{r}
# Define a new variable "num" that adds up the number of characters in the area label

classes |>
  mutate(num = str_length(area))
```

```{r}
# Change the areas to "history", "math", "anthro" instead of "History", "Math", "Anthro"

classes |>
  mutate(area = str_to_lower(area))
```

```{r}
# Create a variable that id's which courses were taught in spring

classes |>
  mutate(spring = str_detect(sem, "SP"))
```

```{r}
# Change the semester labels to "fall2023", "spring2024", "spring2023"

classes |>
  mutate(sem = str_replace(sem, "SP", "spring")) |>
  mutate(sem = str_replace(sem, "FA", "fall"))
```

```{r}
# In the enroll variable, change all e's to 3's (just because?)

classes |>
  mutate(enroll = str_replace_all(enroll, "e", "3"))
```

```{r}
# Use sem to create 2 new variables, one with only the semester (SP/FA) and 1 with the year

classes |>
  mutate(semester = str_sub(sem, 1, 2), 
         year = str_sub(sem, 3, 6))
```

### Summary {-}

Here's what we learned about each function:

-   `str_replace(x, pattern, replacement)` finds the *first* part of `x` that matches the `pattern` and replaces it with `replacement`

-   `str_replace_all(x, pattern, replacement)` finds *all* instances in `x` that matches the `pattern` and replaces it with `replacement`

-   `str_to_lower(x)` converts all upper case letters in `x` to lower case

-   `str_sub(x, start, end)` only keeps a subset of characters in `x`, from `start` (a number indexing the first letter to keep) to `end` (a number indexing the last letter to keep)

-   `str_length(x)` records the number of characters in `x`

-   `str_detect(x, pattern)` is TRUE if `x` contains the given `pattern` and FALSE otherwise

### Example 2

Suppose we *only* want the spring courses:

```{r}
# How can we do this after mutating?
classes |> 
  mutate(spring = str_detect(sem, "SP"))
```

```{r}
# We don't have to mutate first!
classes |> 
  filter(str_detect(sem, "SP"))
```

```{r}
# Yet another way
classes |> 
  filter(!str_detect(sem, "FA"))
```

### Example 3

Suppose we wanted to get separate columns for the first and last names of each course instructor in `classes`. Try doing this using `str_sub()`. But don't try too long! Explain what trouble you ran into.

```{r}
classes |>
  mutate(str_sub(instructor, 1, 4)) |>
  mutate(str_sub(instructor, 5, 20))
```
This is difficult because the instructors' names have varying lengths. 


### Example 4

In general, when we want to split a column into 2+ new columns, we can often use `separate()`:

```{r}
classes |> 
  separate(instructor, c("first", "last"), sep = " ")
```

```{r}
# Sometimes the function can "intuit" how we want to separate the variable
classes |> 
  separate(instructor, c("first", "last"))
```

a.  Separate enroll into 2 separate columns: `students` and `people`. (These columns don't make sense this is just practice).

```{r}
 classes |> 
   separate(enroll, c("students", "people"), sep = "-")
```

b.  We separated `sem` into semester and year above using `str_sub()`. Why would this be hard using `separate()`?

- This would be difficult because there's no character that could separate them with the 'separate()' function. 

c.  When we want to split a column into 2+ new columns (or do other types of string processing), but there's no consistent pattern by which to do this, we can use *regular expressions* (an optional topic):

```{r}
# (?<=[SP|FA]): any character *before* the split point is a "SP" or "FA"
# (?=2): the first character *after* the split point is a 2
classes |> 
  separate(sem, 
          c("semester", "year"),
          "(?<=[SP|FA])(?=2)")
```

```{r}
# More general:
# (?<=[a-zA-Z]): any character *before* the split point is a lower or upper case letter
# (?=[0-9]): the first character *after* the split point is number
classes |> 
  separate(sem, 
          c("semester", "year"),
          "(?<=[A-Z])(?=[0-9])")
```



## Exercises

### Exercise 1: Time slots {.unnumbered}

The `courses` data includes *actual* data scraped from Mac's [class schedule](https://macadmsys.macalester.edu/macssb/customPage/page/classSchedule). (Thanks to Prof Leslie Myint for the scraping code!!)

If you want to learn how to scrape data, take COMP/STAT 212, Intermediate Data Science! NOTE: For simplicity, I removed classes that had "TBA" for the `days`.

```{r}
courses <- read.csv("https://mac-stat.github.io/data/registrar.csv")

# Check it out
head(courses)
```

Use our more familiar wrangling tools to warm up.

```{r}
# Construct a table that indicates the number of classes offered in each day/time slot
# Print only the 6 most popular time slots

courses |>
  count(days,time) |>
  arrange(desc(n)) |>
  head()
```



### Exercise 2: Prep the data {.unnumbered}

So that we can analyze it later, we want to wrangle the `courses` data:

-   Let's get some enrollment info:
    -   Split `avail_max` into 2 separate variables: `avail` and `max`.
    -   Use `avail` and `max` to define a new variable called `enroll`. HINT: You'll need `as.numeric()`
-   Split the course `number` into 3 separate variables: `dept`, `number`, and `section`. HINT: You can use `separate()` to split a variable into 3, not just 2 new variables.

Store this as `courses_clean` so that you can use it later.

```{r}
courses_clean <- courses |>
  separate(avail_max, c("avail", "max"), sep = "/") |>
  mutate(enroll = as.numeric(max) - as.numeric(avail)) |>
  separate(number, c ("dept", "number", "section"))

head(courses_clean)
```



### Exercise 3: Courses by department {.unnumbered}

Using `courses_clean`...

```{r}
# Identify the 6 departments that offered the most sections
courses_clean |>
  count(dept) |>
  arrange(desc(n)) |>
  head()

# Identify the 6 departments with the longest average course titles

courses_clean |>
  mutate(length = str_length(name)) |>
  group_by(dept) |>
  summarize(avg_length = mean(length)) |>
  arrange(desc(avg_length)) |>
  head()
```



### Exercise 4: STAT courses {.unnumbered}

#### Part a {.unnumbered}

Get a subset of `courses_clean` that only includes courses taught by Alicia Johnson.

```{r}
courses_clean |>
  filter(str_detect(instructor, "Alicia Johnson"))
```

#### Part b {.unnumbered}

Create a new dataset from `courses_clean`, named `stat`, that only includes STAT sections. In this dataset:

-   In the course names:

    -   Remove "Introduction to" from any name.
    -   Shorten "Statistical" to "Stat" where relevant.

-   Define a variable that records the `start_time` for the course.

-   Keep only the `number, name, start_time, enroll` columns.

-   The result should have 19 rows and 4 columns.

```{r}
stat <- courses_clean |>
  filter(str_detect(dept, "STAT")) |>
  mutate(name = str_replace(name, "Introduction to ", "")) |>
  mutate(name = str_replace(name, "Statistical", "Stat")) |>
  mutate(start_time = str_sub(time,1,5)) |>
  select(number, name, start_time, enroll)

head(stat)
dim(stat)
```



### Exercise 5: More cleaning {.unnumbered}

In the next exercises, we'll dig into enrollments. Let's get the data ready for that analysis here. Make the following changes to the `courses_clean` data. Because they have different enrollment structures, and we don't want to compare apples and oranges, *remove* the following:

-   all sections in `PE` and `INTD` (interdisciplinary studies courses)

-   all music ensembles and dance practicums, i.e. all MUSI and THDA classes with numbers less than 100. HINT: `!(dept == "MUSI" & as.numeric(number) < 100)`

-   all lab sections. Be careful which variable you use here. For example, you don't want to search by "Lab" and accidentally eliminate courses with words such as "Labor".

Save the results as `enrollments` (don't overwrite `courses_clean`).

```{r}
enrollments <- courses_clean |> 
  filter(dept != "PE", dept != "INTD") |> 
  filter(!(dept == "MUSI" & as.numeric(number) < 100)) |> 
  filter(!(dept == "THDA" & as.numeric(number) < 100)) |> 
  filter(!str_detect(section, "L"))
  
head(enrollments)
```



### Exercise 6: Enrollment & departments {.unnumbered}

Explore enrollments by department. *You* decide what research questions to focus on. Use both visual and numerical summaries.

```{r}
#| fig-height: 10
#| fig-width: 8
ggplot(enrollments, aes(x = enroll, y = dept)) +
  geom_boxplot()
```


### Exercise 7: Enrollment & faculty {.unnumbered}

Let's now explore enrollments by instructor. In doing so, we have to be cautious of cross-listed courses that are listed under multiple different departments. Uncomment the code lines in the chunk below for an example.

::: {.callout-tip title="Commenting/Uncommenting Code"}

To comment/uncomment several lines of code at once, highlight them then click <kbd>ctrl/cmd+shift+c</kbd>.

:::

```{r}
enrollments |>
  filter(dept %in% c("STAT", "COMP"), number == 112, section == "01")
```

Notice that these are the exact same section! In order to not double count an instructor's enrollments, we can keep only the courses that have `distinct()` *combinations* of `days, time, instructor` values.  Uncomment the code lines in the chunk below.

```{r}
enrollments_2 <- enrollments |>
  distinct(days, time, instructor, .keep_all = TRUE)

# NOTE: By default this keeps the first department alphabetically
# That's fine because we won't use this to analyze department enrollments!
 enrollments_2 |> 
   filter(instructor == "Brianna Heggeseth", name == "Introduction to Data Science")
```

*Now*, explore enrollments by instructor. *You* decide what research questions to focus on. Use both visual and numerical summaries.

CAVEAT: The above code doesn't deal with *co-taught* courses that have more than one instructor. Thus instructors that co-taught are recorded as a pair, and their co-taught enrollments aren't added to their total enrollments. This is tough to get around with how the data were scraped as the instructor names are smushed together, not separated by a comma!

```{r}
ggplot(enrollments_2, aes(y = dept, x = enroll, fill = instructor)) + 
  geom_boxplot()
```
