---
title: "Notes"
---

# 4.1 Review - Bivariate Viz 

```{r}
# Import data
survey <- read.csv("https://ajohns24.github.io/data/112/about_us_2024.csv")

# How many students have now filled out the survey? 
dim(survey)
#28 students. 

# What type of variables do we have? 
str(survey)
#Character, integer, numeric, and another character. 
```
**Example 1**: Hangout Preferences 
```{r}
library(tidyverse)

ggplot(survey, aes(x = hangout)) + 
  geom_bar(color = "white", fill = "pink") + 
  labs(x = "Hangout Spot", y = "Number of Students")
```

There was some variability in answers - most students preferred a beach as a hangout spot, and the least amount of students favored a prairie. 

**Example 2**: Temperature Preferences

```{r}
#Plot 1 
ggplot(survey, aes(x = fave_temp)) + 
  geom_histogram(color = "white", binwidth = 15) + 
  labs(x = "Favorite Temperature (Celsius)", y = "Number of Students")

#Plot 2 
ggplot(survey, aes(x = fave_temp)) + 
  geom_density(color = "purple") + 
  labs(x = "Favorite Temperature (Celsius)", y = "Number of Students")
```
Responses were typically around 20 degrees Celsius, with responses typically in the range of 0 degrees to 30 degrees. There is an outlier at about 85 degrees. 

# 4.2 New Stuff 
**Exploring relationships** 
```{r}
weather <- data.frame(temp_3pm = c(24, 26, 20, 15, 15, 0, 40, 60, 57, 44, 51, 75),
                      location = rep(c("A", "B"), each = 6))

weather

ggplot(weather, aes(x = temp_3pm)) +
      geom_density()
```
```{r}
weather <- data.frame(rain_today = c("no", "no", "no", "no", "yes", "no", "yes", "no", "yes", "yes", "no", "yes"),
                        location = c(rep("A", 7), rep("B", 5)))
    weather
    
ggplot(weather, aes(x = location)) +
      geom_bar()
```
# 5.1 Review
```{r}
library(tidyverse)

# Import data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  

# Check out the first 6 rows
head(weather)

# What are the units of observation? -> 2367
nrow(weather)

# How many data points do we have? -> 2367
dim(weather)

# What type of variables do we have?
str(weather)
#We have integer, numeric, and character variables in this data set. 
```
**Example 1** 
```{r}
ggplot(weather, aes (x = temp3pm)) + 
  geom_density()
```
**Example 2** 
Construct 3 plots that address the following research question:

How do afternoon temperatures (temp3pm) differ by location?

```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5)
```
```{r}
# Plot 2 (no facets or densities)
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_histogram()
```

```{r}
# Plot 3 (facets)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density(alpha = 0.5) + 
  facet_wrap(~location)
```

**Example 3** 
```{r}
# Don't worry about the syntax (we'll learn it soon)
woll <- weather |>
  filter(location == "Wollongong") |> 
  mutate(date = as.Date(date))  
```

```{r}
# How often does it raintoday?
# Fill your geometric layer with the color blue.
ggplot(woll, aes(x = raintoday)) + 
  geom_density(fill = "blue")
```
(COME BACK TO)


