---
title: "Univariate Viz"
---

```{r}
library(tidyverse)
```


```{r}
# Import data
hikes <- read.csv("https://mac-stat.github.io/data/high_peaks.csv")
```

# Exercise 1: Research Questions 

```{r}
head(hikes)
```
a. A visualization would help us understand the scales and typical outcomes of the rating variable, and any patterns/relationships that would be helpful to see. 
b. A visualization could help us spot any outliers and patterns and relationships. 

#Exercise 3: Bar Chart of Ratings - Part 1 

```{r}
ggplot(hikes,aes(x=rating))
```
- It returned a bar graph, but since there's no y-axis yet there are no bars. 
- hikes, aes 
- It sets the x-axis as the rating variable. 
- aes might stand aesthetic. 

# Exercise 4: Bar Chart of Ratings - Part 2 

```{r}
# geom_bar () adds the actual visualizations necessary for the bar graph. 
ggplot(hikes, aes(x = rating)) +
  geom_bar()
```
```{r}
# labs(x = "", y = "") adds labels to the x and y axes. 
ggplot(hikes, aes(x = rating)) +
  geom_bar() +
  labs(x = "Rating", y = "Number of hikes")
```
```{r}
# fill = "blue" makes the bar graphs blue! 
ggplot(hikes, aes(x = rating)) +
  geom_bar(fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")
```
```{r}
# color = "orange" makes the border of the bars orange. 
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")
```

```{r}
# theme_minimal() makes the background white. 
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Rating", y = "Number of hikes") +
  theme_minimal()
```

# Exercise 5: Bar Chart Follow-up 

Part a
- The + makes sure the additional lines of code know they are adding onto the original bar graph. 
- "geom" stands for geometric, and makes the geometric shapes we need. 
- labs() stand for labels. 
- Color changes the outline of the bars, while fill fills them in. 

Part b 
- We observed the rating variable, looking at the categories difficult, easy, and moderate. 
- We can now see that most of the hikes are moderate difficulty while the least amount of hikes are difficult. 

Part c 
- I wish that the categories were listed in order from easy to difficult. 

# Exercise 6: Sad Bar Chart 

```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Elevation", y = "Number of hikes") +
  theme_minimal()
```
Since it's quantitative, there are so many varying numbers, so a bar graph visual that shows the mode of each output is not helpful. 

# Exercise 7: A Histogram of Elevation 

Part a 
- About 6. 
- About 2. 

Part b 
- The center of the data points is about 4500. The most typical elevation is between about 3900 to 4100. 
- The min outcome is around 3500. The max is 5500. 
- The range is between 3500 to 5500. It is skewed right. - There are no noticeable outliers, although the 5500 value could be questioned. 

# Exercise 8: Building Histograms - Part 1

```{r}
ggplot(hikes, aes(x = elevation)) + 
  geom_histogram(color = "orange", fill = "blue")  +
  labs(x = "Elevation", y = "Number of hikes") +
  theme_minimal()
```

# Exercise 9: Building Histograms - Part 2 
```{r}
# geom_histogram() outputs a histogram. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram()
```

```{r}
# color = "white" makes the outline of each bin white. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") 
```

```{r}
# fill = "blue" makes the full color of each bin blue. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", fill = "blue") 
```

```{r}
# Adds proper labels to the axes. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white") +
  labs(x = "Elevation (feet)", y = "Number of hikes")
```

```{r}
# It sets the range of the elevation for each bin. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 1000) +
  labs(x = "Elevation (feet)", y = "Number of hikes")
```

```{r}
# This made the bins much smaller, since each only contains 5 feet. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 5) +
  labs(x = "Elevation (feet)", y = "Number of hikes")
```

```{r}
# A binwidth of 200 appears to be the most appropriate here, as there are enough bins for us to see a pattern in the data without it feeling too crowded. 
ggplot(hikes, aes(x = elevation)) +
  geom_histogram(color = "white", binwidth = 200) +
  labs(x = "Elevation (feet)", y = "Number of hikes")
```
# Exercise 10: Histogram Follow-up

- geom_histogram()
- "color" changes the border of the bins, while "fill" changes the color of the bins themselves. 
- Adding color="white" makes it easier to see the difference between each bin, as that can be hard to see especially in a histogram. 
- "binwidth" changes the width of each bin, or the values of the x variable that the data points will fall into. 
- If the binwidth is too big, then the histogram isn't informative as the bins are too large to see any patterns in the data. 
- If the binwidth is too small, then the histogram isn't informative as it becomes impossible to see trends in the data. 

# Exercise 11: Density Plots 

```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_density()
```
- I believe that adding geom_density(color = "blue") will make the line blue and that adding geom_density(fill = "orange") will make the area underneath the line orange.  

```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_density(color = "blue", fill = "orange")
```
- The density plot makes it easier to see the peak of observations based on their elevation. The hikes' elevation typically tends to be a bit above 4000 feet. The data is skewed right and seem to be between 3500 feet to about 6000 feet. 

# Exercise 12: Density Plots vs Histograms
- Since histograms collect the data into bins, it might not be ideal for catching subtle behaviors of the data in comparison to a density plot, which smooths out the data. However, histograms are easier to interpret and might be more effective for small data sets. 

# Exercise 13: Code = communication 
General structure of these codes: 
ggplot(___, aes(x = ___)) + 
  geom___(color = "___", fill = "___") + 
  labs(x = "___", y = "___")
# Exercise 14: Practice 
Part a 

```{r}
# Data on students in this class
survey <- read.csv("https://hash-mac.github.io/stat112site-s25/data/survey.csv")

ggplot(survey, aes(x = hangout)) + 
  geom_bar(color = "white", fill = "salmon") +
  labs(x = "Favorite Hangout Spot", y = "Number of Students")
```
