---
title: "Bivariate Viz"
---

```{r}
library(tidyverse)
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```
# Exercise 0: Review 
**Part a** 
How many, or roughly what percent, of the 3000+ counties did the Republican candidate win in 2020?
- My guess: 72 % 

Then make a plot of the winner variable.
```{r}
ggplot(elections, aes(x = winner_20)) + 
  geom_bar(color = "white", fill = "salmon") + 
  labs(x = "2020 Winner", y = "Number of Counties")
```

Follow-up Questions: 
- Overall winner? 
- Still need the percentage. 
- Number of counties per state? 

**Part b** 
```{r}
ggplot(elections, aes(x = repub_pct_20)) + 
  geom_histogram(color = "seashell", fill = "peachpuff") + 
  labs(x = "Percentage of Republican Support", y = "Number of Counties")
```
- What is the overall percentage of Republican voters? 
- What does the percentage of Democratic support look like? 
- How many counties are within each state? What is their weight to the overall vote? 

# Exercise 1: Quantitative vs Quantitative Intuition Check 
```{r}
ggplot(elections, aes(x = repub_pct_16, y = repub_pct_20)) +
  geom_point()
```
# Exercise 2: 2 Quantitative Variables 
```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone? It is a bivariate visualization instead of a univariate, as it is reflecting the relationship between two variables rather than one. 
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16))
```
```{r}
# Add a layer of points for each county
# Take note of the geom! -> geom_point()
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point()
```
```{r}
# Change the shape of the points
# What happens if you change the shape to another number? -> It makes the shape of each point different. In this case, they look like pluses rather than dots. 
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(shape = 3)
```

```{r}
# YOU TRY: Modify the code to make the points "orange"
# NOTE: Try to anticipate if "color" or "fill" will be useful here. Then try both.
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = "plum", fill = "palegreen", shape = 3)
#"Color" is more useful.  
```
```{r}
# Add a layer that represents each county by the state it's in
# Take note of the geom and the info it needs to run!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_text(aes(label = state_abbr))
#"geom_text" and "label = variable". 
```
# Exercise 3: Reflect 
- Republican support in 2016 has a strong positive relationship with republican support in 2020. There are no noticeable outliers, as all the points reflect a near linear line. 

# Exercise 4: Visualizing trend 
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth()
#Adds a fitted line. 
```
**Part a** 
```{r}
#Contains the model smooth without the individual point glyphs. 
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_smooth()
```

**Part b** 
```{r}
#Adds a linear line of best fit. 
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth(method = "lm")
```
# Exercise 5: Your Turn 
```{r}
# Scatterplot of repub_pct_20 vs median_rent
ggplot(elections, aes(y = repub_pct_20, x = median_rent)) +
  geom_point() +
  geom_smooth(method = "lm")

# Scatterplot of repub_pct_20 vs median_age
ggplot(elections, aes(y = repub_pct_20, x = median_age)) +
  geom_point() +
  geom_smooth(method = "lm")
```
- The relationship between median_rent and repub_pct_20 is a moderate to weak negative correlation. 
- The relationship between median_age and repub_pct_20 is a weak positive correlation. 
- It is hard to tell without a clear linear regression analysis which variable is a better predictor of repub_pct_20, but from my eyes I would predict that median_rent is a slighter better predictor as it seems to follow a more clear pattern than median_age. 

# Exercise 6: A Sad Scatterplot 
```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_point()
```
Since "historical" is categorical, it tells us a count of observations in each category, which does not work well against a quantitative variable as its scatterplot reveals no information about their relationship with each other. We don't know the relationship between the Republican voting trends of 2020 and historical voting trends. 

# Exercise 7: Quantitative vs Categorical - Violins & Boxes 
```{r}
# Side-by-side violin plots
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_violin()
```

```{r}
# Side-by-side boxplots (defined below)
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_boxplot()
#Box plots are constructed from five numbers - the minimum, 25th percentile, median, 75th percentile, and maximum value of a quantitative variable. 
```
**Reflect** 
Historically, counties with a higher republican voting percentage in the 2020 election voted red. Then, they were second most likely to have historically voted purple, and least likely to have voted blue. 

# Exercise 8: Quantitative vs Categorical - Intuition Check 

```{r}
ggplot(elections, aes(x = repub_pct_20)) +
  geom_density(fill = "blue")
```
# Exercise 9: Quantitative vs Categorical - Density Plots 
```{r}
# Name two "bad" things about this plot: 1) There are no labels; 2) The colors don't match the parties. 
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```
```{r}
# What does scale_fill_manual do? -> It changes the color of each category.  
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
```{r}
# What does alpha = 0.5 do? -> It adjusts the transparency
# Play around with different values of alpha, between 0 and 1
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
```{r}
# What does facet_wrap do?! -> facet_wrap breaks the graph into its different categories for easy individual viewing of each category. 
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red")) +
  facet_wrap(~ historical)
```

```{r}
# Let's try a similar grouping strategy with a histogram instead of density plot.
# Why is this terrible? -> Since histograms prioritize the count over the ratio of each category, the visualization becomes incredibly misleading, as blue is now the seemingly most popular party historically. 
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_histogram(color = "white") +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

# Exercise 10 
- I like density plots the most because it seems to be most efficient for immediate pattern-recognition. The others take a bit more effort in order to understand the distribution of each the categories, although I like boxplots as well. 
- Density plots tell us more details about the distribution of the data because it is easy to see the shape and modes. 
- One con, however, is that density plots can be harder to compare details about the relationships between variables, while boxplots are a great device for visually comparing percentiles, medians, etc. 

# Exercise 11: Categorical vs Categorical - Intuition Check 
```{r}
# Plot 1: adjust this to recreate the top plot
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```
```{r}
# Plot 2: adjust this to recreate the bottom plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```
# Exercise 12: Categorical vs Categorical 

```{r}
# A stacked bar plot
# How are the "historical" and "winner_20" variables mapped to the plot, i.e. what roles do they play?
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
#We are measuring the count of each county historically, moderated by the how many counties each party one in 2020. 
```
```{r}
# A faceted bar plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```

```{r}
# A side-by-side bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```
```{r}
# A proportional bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```
Part a 
- It is not helpful for understanding the total number of counts for each category, but
- It is helpful for understanding the percentage each variable received. 

Part b 
- I think it depends on the question. I think the side-by-side bar plot is the most applicable to the majority of scenarios, but if I needed to understand percentages the proportional bar plot would be more handy. 

# Exercise 13: Practice 

```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv")
```

**How do 3pm temperatures (temp3pm) differ by location?** 
```{r}
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_boxplot()
```
Uluru seems to have the most varying temperatures, but overall the highest at a median temperature of 30 degrees. Hobart and Wollongong are close, as the lowest 25th percentile of temperatures in Wollongong overlaps with the highest 25th percentile of Hobart, but overall Hobart has a colder median temperature of about 15 degrees while Wollongong is closer to 20. 

**# How might we predict the 3pm temperature (temp3pm) by the 9am temperature (temp9am)?** 
```{r}
ggplot(weather, aes(x = temp9am, y = temp3pm)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```
Overall, the temperature at 3 pm seems to be consistently 5 degrees higher than it was at 9 am. 

**# How do the number of rainy days (raintoday) differ by location?**
```{r}
ggplot(weather, aes(fill = raintoday, x = location)) + 
  geom_bar()
```
None of the locations seem particularly rainy, but Uluru is certainly the least rainy and Wollongong seems to be the relatively most rainy. 

# 4.4 Exercises (opt)
